---
title: "Semi-Supervised STACAS demo"
author: "A. Berenstein, C.Amezcua"
date: "22/07/2022"
output:
  rmdformats::readthedown:
    self-contained: true
    highlight: haddock
    thumbnails: false
    css: styles.css
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file, encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'PanBcellDef.html'))})
---

# Introduction 
This notebook will work on data produced by Amezcua et al. (Cell 2019)
In that work, Amezcua analyzed tissue resident Th cells. 
Now, the idea is to recycle those data but redirecting the focus on Bcells. In particular, those tissue resident ones. 


```{r}
renv::activate()
renv::restore()
renv::snapshot()
#install.packages("Seurat")
#install.packages("remotes")
#remotes::install_github("carmonalab/UCell")
#remotes::install_github("carmonalab/scGate")
#install.packages("BiocManager")
#BiocManager::install("SingleR")
#BiocManager::install("celldex")
#BiocManager::install("ontoProc")
#BiocManager::install("rmarkdown")

library("Seurat")
library("dplyr")
library("UCell")
library("scGate")
library("SingleR")
library("ggplot2")
source("./utils.R")

```

# Load datasets from counts
```{r}
path <- c("~/Dropbox/Amezcua/data")
samples <- c("DS-017-WT","DS-017-Im1","DS-017-Im2")
count_files <- paste(path,samples,"gene_exon_tagged_dge.txt.gz",sep = "/")

# Output files:
panBellFile <-  "~/Dropbox/Amezcua/processed_data/seurat_processed_07_22.rds"  # Subseted Pan Bcell FIle with recomputed UMAP

```

## Load proccessed object
```{r}
if(file.exists(panBellFile)){
  run =F
  seurat.merged.processed <- seurat.pan.Bcells <- readRDS(panBellFile)  
}else{
  run =T
}

```


#create Seurat objects
```{r}
if(run){
  seuratDataList <- list()
  for(i in seq_along(samples)){
    countdata <- read.table(
      file = count_files[i],
      header = T,
      row.names = 1
    )
      
    obj <- Seurat::CreateSeuratObject(counts = countdata,project = samples[i])
    seuratDataList[[i]] <- obj
  }
  ncell <- table(unlist(sapply(seuratDataList,colnames)))
  to.drop <- names(ncell[ncell>1])
  
  # remove 4 barcodes duplicated across samples
  seuratDataList = lapply(seuratDataList,function(x){
    res = subset(x,cells = setdiff(colnames(x),to.drop) )
    return(res)
  })
}
```


# quality filtering
```{r,fig.height= 8,fig.width= 13}
if(run){
  seurat.merged <- Reduce(merge, seuratDataList)
  seurat.merged@meta.data%>%head()
  
  percent.ribo.dv <- PercentageFeatureSet(seurat.merged, pattern = "^Rp[ls]")
  percent.mito.dv <- PercentageFeatureSet(seurat.merged, pattern = "^mt-")
  
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.ribo.dv, col.name = "percent.ribo")
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.mito.dv, col.name = "percent.mito")
Seurat::VlnPlot(seurat.merged,features = c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),group.by = "orig.ident",ncol = 2)
}  
```

```{r}
splited.by.sample <- SplitObject(seurat.merged, split.by = "orig.ident")
x <- splited.by.sample[[3]]
q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
q
```

# Filter low or high expressed cells 
```{r}
if(run){
  splited.by.sample <- SplitObject(seurat.merged, split.by = "orig.ident")
  
  splited.by.sample.qc <- lapply(splited.by.sample, function(x) {
     #n UMIs
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nCount_RNA>max(q["10%"]) & nCount_RNA<q['95%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nFeature_RNA>max(q['10%']) & nFeature_RNA < q['95%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.ribo>q['1%'] & percent.ribo < min(50,q['99%']) )
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.mito < q['99%'])
     return(x)
  })
  
  seurat.filtered.merged <- Reduce(merge, splited.by.sample.qc)
  dim(seurat.merged)
  dim(seurat.filtered.merged)
}else{
  seurat.filtered.merged <- seurat.merged.processed
}

sapply(c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),function(x){summary(seurat.filtered.merged[[x]])})
seurat.filtered.merged
```

## Control step: how many cells were dropped by high expression values? (Re: 262 cells)
```{r}
if(run){
  high.expr.dropped <- lapply(splited.by.sample, function(x){
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_ncount <- subset(x, nCount_RNA>q['95%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_nfeature <- subset(x, subset= nFeature_RNA > q['95%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_ribo <- subset(x, subset=percent.ribo > min(50,q['99%']))
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_mito <- subset(x, subset=percent.mito > q['99%'])
     
     return(table(c(x_ncount%>%colnames(), x_nfeature%>%colnames(),x_ribo%>%colnames(),x_ribo%>%colnames())))
    }
  )
  
  # 262 cells dropped by staying over the top 1% quantile in any of control quantity
  cells.dropped.by.hihg.quantile <- unlist(high.expr.dropped)
  cells.dropped.by.hihg.quantile%>%length()
}
```


## Sanity check: visual detection of possible bach effects (Re: thats ok)
```{r,fig.height= 4.5,fig.width= 12}
ndim = 20      # for pca
nfeat = 1000    # number of most variable features to be considered

# scale data; find variable features, run pca and run umap
if(run) seurat.merged.processed <- process_obj(seurat.filtered.merged,ndim = ndim, nfeat = nfeat, find_variable_features = T) 
DimPlot(seurat.merged.processed,,split.by = "orig.ident")
```

# 7275 cells
# 18.886 features 
```{r}
seurat.merged.processed
```

## run singleR
```{r}
if(run){
  ref.se <- celldex::ImmGenData()
  param <- BiocParallel::MulticoreParam(workers = 32)
  
  ## broad 
  result.singler.main = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se, labels = ref.se$label.main, BPPARAM=param)
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler.main$pruned.labels,col.name = "SingleR.main")

  if(F){  
  # finer (ontology based)
    result.singler = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se,labels = ref.se$label.ont, BPPARAM=param)
    # get term names
    cl <- ontoProc::getCellOnto()
    terms <- result.singler$pruned.labels%>%unique()
    getTerms <- ontoProc::label_TAG(ontology = cl,terms)
    seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler$pruned.labels,col.name = "SingleR.tag")
    seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = unname(getTerms[result.singler$pruned.labels]),col.name = "SingleR.term")
  }
}

```

How many LinB cells do we filtered out? 
```{r}
(seurat.merged.processed%>%subset(SingleR.main == "B cells"))$orig.ident%>%table()
```


# visualize SingleR classification

```{r,fig.height= 6,fig.width= 6}
Fig1a = DimPlot(seurat.merged.processed,group.by = "SingleR.main",label =T,repel =T,label.size = 3) + theme(aspect.ratio = 1) + labs(tag = "A")
ggsave("plots/Jul22/fig1a_singleR.pdf",plot = Fig1a,width = 7, height = 7)
ggsave("plots/Jul22/fig1a_singleR.png",plot = Fig1a,width = 7, height = 7)
ggsave("plots/Jul22/fig1a_singleR.tiff",plot = Fig1a,width = 7, height = 7)
Fig1a

```

# Clustering whole assay
```{r}
if(run){
  set.seed(1234)
  seurat.merged.processed <- FindNeighbors(seurat.merged.processed, dims = 1:20)
  seurat.merged.processed <- FindClusters(seurat.merged.processed, resolution = 0.7)
}
```


```{r,fig.height= 5,fig.width= 5}
Fig1b  <- DimPlot(seurat.merged.processed,group.by = "seurat_clusters",label = T,label.size = 3) + theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + labs(tag = "B")
ggsave("plots/Jul22/Fig1b.pdf", plot = Fig1b,width = 7, height = 7)
ggsave("plots/Jul22/Fig1b.png",plot = Fig1b,width = 7, height = 7)
ggsave("plots/Jul22/Fig1b.tiff",plot = Fig1b,width = 7, height = 7)
bcell_clusters = c("0","2","10","17")
Fig1b

```


```{r,fig.height= 7,fig.width= 9}

SFig1a <- DimPlot(seurat.merged.processed%>%subset(SingleR.main == "B cells"), group.by = "SingleR.main",label =T,repel =T,label.size = 3) +
  theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("SingleR B cells") + labs(tag = "A")
SFig1a

SFig1b <- DimPlot(seurat.merged.processed%>%subset(seurat_clusters %in% bcell_clusters),group.by = "seurat_clusters",label = T,label.size = 3) + 
  NoLegend() +  theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("Only BCell clusters (0,2,10,17)") + labs(tag = "B")

notBcell = setdiff(seurat.merged.processed$seurat_clusters%>%unique() ,bcell_clusters) 
SFig1c <- DimPlot(seurat.merged.processed%>%subset(seurat_clusters %in% notBcell),group.by = "seurat_clusters",label = T,label.size = 3)  + 
  NoLegend() +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("Removing BCell clusters") + labs(tag = "C")

SFig1d <- DimPlot(seurat.merged.processed%>%subset(seurat_clusters %in% "7"),group.by = "seurat_clusters",label = T,label.size = 3) + 
  NoLegend() +  theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + ggtitle("Mac-Mono Cluster (7)") + labs(tag = "D")



SFig1 <- SFig1a + SFig1b + SFig1c + SFig1d

ggsave("plots/Jul22/SFig1_singleR.pdf",plot = SFig1,width = 12, height = 10)
ggsave("plots/Jul22/SFig1_singleR.png",plot = SFig1,width = 12, height = 10)
ggsave("plots/Jul22/SFig1_singleR.tiff",plot = SFig1,width = 12, height = 10)
SFig1

```



# scGate
The idea is to filter out some impure cells/doublets that could be passed thorough SingleR filtering model
```{r, fig.height= 4, fig.width= 5}
panBcell <- scGate::gating_model(name = "cd79", signature = "Cd79a")
panBcell <- scGate::gating_model(panBcell,name = "cd19", signature = "Cd19")
panBcell <- scGate::gating_model(panBcell, name = "cd20", signature = "Ms4a1")
panBcell <- scGate::gating_model(panBcell, name = "plasmaticas_residentes", signature = "Sdc1;Igha")
panBcell <- scGate::gating_model(panBcell, name = "Tcell", signature = "Cd3e;Cd3d;Cd4;Cd8a", negative = T)
panBcell <- scGate::gating_model(panBcell, name = "myeloid", signature = "Spi1", negative = T)
panBcell <- scGate::gating_model(panBcell, name = "MoMacDC", signature = "Lyz1;Lyz2;Csf1r;Msr1;Mafb;Cd300e", negative = T)
panBcell <- scGate::gating_model(panBcell, name = "Macrophage", signature = "Cd40;Cd68;Cd14;Fcgr1a", negative = T)
```

```{r, fig.height= 4.5,fig.width= 5}
set.seed(1234)

seurat.merged.processed <- scGate(data = seurat.merged.processed, model = panBcell, output.col.name = "panBcell")

Fig1c <- DimPlot(seurat.merged.processed, cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell" ) + 
  theme(aspect.ratio = 1) + ggtitle("scGate: panBcell model") + theme(aspect.ratio = 1)  +  xlim(-20, 15) + ylim(-12, 17) + labs(tag = "C")
seurat.merged.processed$panBcell%>%table()

ggsave("plots/Jul22/Fig1c.pdf",plot = Fig1c,width = 7, height = 7)
ggsave("plots/Jul22/Fig1c.png",plot = Fig1c,width = 7, height = 7)
ggsave("plots/Jul22/Fig1c.tiff",plot = Fig1c,width = 7, height = 7)

Fig1c

```



## Filtering Bcells
```{r}
seurat.merged.processed$SinlgeR_Seurat_scGate = (seurat.merged.processed$SingleR.main == "B cells") & (seurat.merged.processed$seurat_clusters %in% bcell_clusters) & (seurat.merged.processed$panBcell == "Pure") 

seurat.merged.processed$SinlgeR_Seurat_scGate%>%table()
(seurat.merged.processed%>%subset(SinlgeR_Seurat_scGate == TRUE))$orig.ident%>%table()  

seurat.pan.Bcells <- seurat.merged.processed %>%subset(SinlgeR_Seurat_scGate == TRUE)
seurat.pan.Bcells
```


```{r}
SFig2 <- DimPlot(seurat.pan.Bcells,group.by = "SinlgeR_Seurat_scGate",label.size = 3) + theme(aspect.ratio = 1) +  xlim(-20, 15) + ylim(-12, 17)  + NoLegend() + ggtitle("SingleR + SeuratClusters + scGate")
ggsave("plots/Jul22/SFig2.pdf",plot = SFig2,width = 7, height = 7)
ggsave("plots/Jul22/SFig2.png",plot = SFig2,width = 7, height = 7)
ggsave("plots/Jul22/SFig2.tiff",plot = SFig2,width = 7, height = 7)
SFig2
```

Manually remove outliers
```{r}
aux = DimPlot(seurat.pan.Bcells)
seurat.pan.Bcells$outliers_umap <- (aux$data$UMAP_1 < -5 | aux$data$UMAP_2 < -1)
seurat.pan.Bcells <- seurat.pan.Bcells%>%subset(outliers_umap == FALSE)
seurat.pan.Bcells$orig.ident%>%table()  

Fig1d <- DimPlot(seurat.pan.Bcells,group.by = "SinlgeR_Seurat_scGate",label.size = 3) + theme(aspect.ratio = 1) +  xlim(-20, 15) + ylim(-12, 17)  + NoLegend() + ggtitle("lin B cells")  + labs(tag = "D")

ggsave("plots/Jul22/Fig1d.pdf",plot = Fig1d,width = 7, height = 7)
ggsave("plots/Jul22/Fig1d.png",plot = Fig1d,width = 7, height = 7)
ggsave("plots/Jul22/Fig1d.tiff",plot = Fig1d,width = 7, height = 7)


Fig1d
```

```{r, fig.height= 10,fig.width= 12}
Fig1 <-  Fig1a + Fig1b + Fig1c + Fig1d
ggsave("plots/Jul22/Fig1.pdf",plot = Fig1,width = 12, height = 10)
ggsave("plots/Jul22/Fig1.png",plot = Fig1,width = 12, height = 10)
ggsave("plots/Jul22/Fig1.tiff",plot = Fig1,width = 12, height = 10)
Fig1

```


## Save filtered dataset
```{r}
if(run){
  set.seed(1234)
  saveRDS(seurat.pan.Bcells,file = panBellFile)
}
```

## Split WT and IM datasets and analize them
# Load Bcell-filtered dataset
```{r,warning=F, collapse =T,message =F}
ndim = 30; seed = 31416
data <- readRDS(panBellFile)
data%>%dim
data$condition <- ifelse(data$orig.ident == "DS-017-WT","wt","im") 
data.bkp <- data
data.wt <- data%>%subset(orig.ident == "DS-017-WT")
data.im <- data%>%subset(orig.ident != "DS-017-WT")
```
Visualize datasets by condition
```{r,fig.width=6,fig.height=3,warning=F, collapse =T,message =F}
fig2 <- DimPlot(data,group.by = "orig.ident",split.by = "orig.ident")
ggsave("plots/Jul22/Fig2.pdf",plot = fig2,width = 12, height = 4)
ggsave("plots/Jul22/Fig2.png",plot = fig2,width = 12, height = 4)
ggsave("plots/Jul22/Fig2.tiff",plot = fig2,width = 12, height = 4)
fig2
```

# Preprocessing

**We will preprocess separately WT and IM subsets**

## Cell cycle scoring and regression 

The idea here is to mitigate the effects of cell cycle heterogeneity by calculating cell cycle phase scores based on canonical markers, and regressing these out of the data during pre-processing (see [SeuratVignette cell cycle](https://satijalab.org/seurat/articles/cell_cycle_vignette.html) for details)

```{r,warning=F, collapse =T,message =F, results='hide'}
# extract cell cycle genes from Seurat 
s.genes <- cc.genes$s.genes%>%hs.2.mm(.)
g2m.genes <- cc.genes$g2m.genes%>%hs.2.mm(.)
```


```{r,warning=F, collapse =T,message =F, results='hide'}
# WT
set.seed(seed)
data.wt <- CellCycleScoring(data.wt, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.wt$CC.Difference <- data.wt$S.Score - data.wt$G2M.Score
data.wt <- SCTransform(data.wt, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE)
data.wt <- RunPCA(data.wt, features = data.wt@assays$RNA@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.wt <- RunUMAP(data.wt, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)

```



```{r,warning=F, collapse =T,message =F,results='hide'}
# IM
set.seed(seed)

data.im <- CellCycleScoring(data.im, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.im$CC.Difference <- data.im$S.Score - data.im$G2M.Score

# work as usual (only variable genes)
data.im <- SCTransform(data.im, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE)
data.im <- RunPCA(data.im, features = data.im@assays$RNA@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.im <- RunUMAP(data.im, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)

```

# Unsupervised analysis
```{r,fig.height=4, fig.width= 10,warning=F, collapse =T,message =F, results='hide'}
set.seed(seed)
data.im <- FindNeighbors(data.im,assay  = "SCT", dims = 1:20)
data.im <- FindClusters(data.im, resolution = 0.7)
data.im <- FindClusters(data.im, resolution = 0.5)
data.im <- FindClusters(data.im, resolution = 0.4)
data.im <- FindClusters(data.im, resolution = 0.3)
data.im <- FindClusters(data.im, resolution = 0.35)


set.seed(seed)
data.wt <- FindNeighbors(data.wt,assay  = "SCT", dims = 1:20)
#data.wt <- FindClusters(data.wt, resolution = 2)
data.wt <- FindClusters(data.wt, resolution = 1.5)
data.wt <- FindClusters(data.wt, resolution = 1)
data.wt <- FindClusters(data.wt, resolution = 0.9)
data.wt <- FindClusters(data.wt, resolution = 0.8)
data.wt <- FindClusters(data.wt, resolution = 0.75)

data.wt <- FindClusters(data.wt, resolution = 0.7)

data.wt <- FindClusters(data.wt, resolution = 0.6)
data.wt <- FindClusters(data.wt, resolution = 0.65)

data.wt <- FindClusters(data.wt, resolution = 0.5)
data.wt <- FindClusters(data.wt, resolution = 0.4)
data.wt <- FindClusters(data.wt, resolution = 0.3)


Fig3a <- DimPlot(data.wt,group.by = "SCT_snn_res.0.8",label = T)
Fig3b <- DimPlot(data.im ,group.by = "SCT_snn_res.0.4",label = T)

Fig3a + Fig3b

ggsave("plots/Jul22/Fig3a_clustering_wt_im.pdf",plot = Fig3a,width = 4.5, height = 4) 
ggsave("plots/Jul22/Fig3a_clustering_wt_im.png",plot = Fig3a,width = 4.5, height = 4)
ggsave("plots/Jul22/Fig3a_clustering_wt_im.tiff",plot = Fig3a,width = 4.5, height = 4)

ggsave("plots/Jul22/Fig3b_clustering_wt_im.pdf",plot = Fig3b,width = 4.5, height = 4)
ggsave("plots/Jul22/Fig3b_clustering_wt_im.png",plot = Fig3b,width = 4.5, height = 4)
ggsave("plots/Jul22/Fig3b_clustering_wt_im.tiff",plot = Fig3b,width = 4.5, height = 4)


```

# relation ids old and new clusters (Para que a Caro se le haga más fácil reinterpretar)
```{r}
test <- data.im
original_clustes = read.table("~/Dropbox/Amezcua/tissue_resident_cells/reports/old_im_clusters_0_to_7.tsv")
colnames(original_clustes) <- c("id","old_SCT_snn_res.0.4")

im_clusters <- test@meta.data%>%select("SCT_snn_res.0.4")
im_clusters$id <- im_clusters%>%rownames()

im_clusters = merge(im_clusters,original_clustes,by.x = "id",by.y = "id",all.x = F,all.y =F)


relation_ids_im = table(im_clusters$SCT_snn_res.0.4,im_clusters$old_SCT_snn_res.0.4)
write.table(relation_ids_im,"~/Dropbox/Amezcua/tissue_resident_cells/reports/cluster_relation_ids_old_new_UMAPs.tsv",quote = F)


test <- data.wt
original_clustes = read.table("~/Dropbox/Amezcua/tissue_resident_cells/reports/old_wt_clusters_0_to_3.tsv")
colnames(original_clustes) <- c("id","old_SCT_snn_res.0.8")

wt_clusters <- test@meta.data%>%select("SCT_snn_res.0.8")
wt_clusters$id <- wt_clusters%>%rownames()

wt_clusters = merge(wt_clusters,original_clustes,by.x = "id",by.y = "id",all.x = F,all.y =F)


relation_ids_wt = table(wt_clusters$SCT_snn_res.0.8,wt_clusters$old_SCT_snn_res.0.8)
relation_ids_wt
write.table(relation_ids_wt,"~/Dropbox/Amezcua/tissue_resident_cells/reports/wt_cluster_relation_ids_old_new_UMAPs.tsv",quote = F)

```

## Write cluster ids to disk
```{r}
write.table(data.im$SCT_snn_res.0.4,"~/Dropbox/Amezcua/tissue_resident_cells/reports/im_jul22_clusters_0_to_7.tsv",sep = "\t",quote = F,col.names = F)
write.table(data.wt$SCT_snn_res.0.8,"~/Dropbox/Amezcua/tissue_resident_cells/reports/wt_jul22_wt_clusters_0_to_2.tsv",sep = "\t",quote = F,col.names = F)
```


## #############################
# Marker analysis

```{r,fig.height= 20,fig.width=10,warning=F, collapse =T,message =F}
Idents(data.wt) <- "SCT_snn_res.0.8"
discrim.wt <- FindAllMarkers(data.wt, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.wt <- discrim.wt%>%subset(p_val_adj < 0.05)
gene.wt <- discrim.wt$gene
d <- DoHeatmap(data.wt,features = gene.wt)

Idents(data.im) <- "SCT_snn_res.0.4"
discrim.im <- FindAllMarkers(data.im,assay = "SCT", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.im.uncorrected <- discrim.im
discrim.im <- discrim.im%>%subset(p_val_adj < 0.05)

gene.im <- discrim.im$gene

e <- DoHeatmap(data.im,features = gene.im,size = 3,lines.width = 35)

# show WT heatmap
d 

# show IM heatmap
e 
```


```{r,eval = F}
ggsave(d,file = "plots/Jul22/markers_heatmaps/heatmap_fdr005.wt.png",height = 20,width = 10, dpi = 300)
ggsave(d,file = "plots/Jul22/markers_heatmaps/heatmap_fdr005.wt.tiff",height = 20,width = 10, dpi = 300)
ggsave(d,file = "plots/Jul22/markers_heatmaps/heatmap_fdr005.wt.pdf",height = 20,width = 10,dpi = 300)
write.table(discrim.wt,file = "plots/Jul22/markers_heatmaps/markers_fdr005.wt.tsv",sep = "\t",quote =F,row.names =F)

ggsave(e,file = "plots/Jul22/markers_heatmaps/heatmap_fdr005.im.png",height = 20,width = 10,dpi = 300)
ggsave(e,file = "plots/Jul22/markers_heatmaps/heatmap_fdr005.im.tiff",height = 20,width = 10,dpi = 300)
ggsave(e,file = "plots/Jul22/markers_heatmaps/heatmap_fdr005.im.pdf",height = 20,width = 10,dpi = 300)
write.table(discrim.im,file = "plots/Jul22/markers_heatmaps/markers_fdr005.im.tsv",sep = "\t",quote =F,row.names =F)
```


```{r,eval = F}
# 
discrim.wt.down <- FindAllMarkers(data.wt, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.wt.down <- discrim.wt.down%>%subset(p_val_adj < 0.05)
write.table(discrim.wt.down,file = "plots/Jul22/markers_heatmaps/markers_fdr005.wt_include_downReg.tsv",sep = "\t",quote =F,row.names =F)


discrim.im.down <- FindAllMarkers(data.im, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.im.down <- discrim.im.down%>%subset(p_val_adj < 0.05)
write.table(discrim.im.down,file = "plots/Jul22/markers_heatmaps/markers_fdr005.im_include_downReg.tsv",sep = "\t",quote =F,row.names =F)


```



```{r,eval=F,echo=F,results=F}
discrim.im$cluster%>%table()
discrim.im%>%head((discrim.im$cluster%>%table())["0"])
discrim.im%>%subset(cluster==5)
```


# DimPlot using Heatmap genes

Plot again clusters in WT and IM datasets


```{r,fig.height=4, fig.width= 10,warning=F, collapse =T,message =F, results='hide'}
a + b
```


WT dataset
```{r,fig.width= 15,fig.height= 50,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = discrim.wt$gene,ncol = 5)
```

IM dataset
```{r,fig.width= 20,fig.height=70,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = intersect(discrim.im$gene,data.im%>%rownames()),ncol = 6)
```


# Explore predefined markers 

We will use here those genes previously defined in our molecular signature spreadsheet file

```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
features1 = c("Ighm","Ighd1-1","Igha","Ighg1","Cd19","Ms4a1","Fcer2a","Cd24a","Cd38","Pdcd1lg2","Nt5e","Entpd1","Cd80","Cd86","Cd93","Cd69","Itgal","Cd44","Cd28","Sdc1",	"Prdm1","Tnfrsf17","Sell","Igj")
features2 <- c("Fas","Bcl6","Cd40","Cxcr5","Cxcr4")
```

```{r,fig.height= 8,fig.width=8,warning=F, collapse =T,message =F}
h1 <- DoHeatmap(data.wt,features = features1)
h2 <- DoHeatmap(data.im,features = features1,size = 3,lines.width = 35)

h1
h2
```


WT dataset


```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = features1, ncol = 5)
```

IM dataset
```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = features1, ncol = 5)
```


WT dataset (new genes)
```{r,fig.height=7,fig.width=12,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = features2, ncol = 3)
```

IM dataset (new genes)
```{r,fig.height=7,fig.width=12,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = features2, ncol = 3)
```


# Explore markers of resident plasmatic cells (IgA+)
WT dataset
```{r,fig.width= 12,fig.height=8,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```

IM dataset
```{r,fig.width= 13,fig.height=7,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```



