---
title: "Semi-Supervised STACAS demo"
author: "A. Berenstein, C.Amezcua"
date: "15/01/2022"
output:
  rmdformats::readthedown:
    self-contained: true
    highlight: haddock
    thumbnails: false
    css: styles.css
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file, encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'PanBcellDef.html'))})
---

# Introduction 
This notebook will work on data produced by Amezcua et al. (Cell 2019)
In that work, Amezcua analyzed tissue resident Th cells. 
Now, the idea is to recycle those data but redirecting the focus on Bcells. In particular, those tissue resident ones. 


```{r}
renv::activate()
renv::restore()
#install.packages("Seurat")
#install.packages("remotes")
#remotes::install_github("carmonalab/UCell")
#remotes::install_github("carmonalab/scGate")
#install.packages("BiocManager")
#BiocManager::install("SingleR")
#BiocManager::install("celldex")
#BiocManager::install("ontoProc")
#BiocManager::install("rmarkdown")

library("Seurat")
library("dplyr")
library("UCell")
library("scGate")
library("SingleR")
library("ggplot2")
source("./utils.R")

```

# Load datasets from counts
```{r}
path <- c("~/Dropbox/Amezcua/data/")
samples <- c("DS-017-WT","DS-017-Im1","DS-017-Im2")
count_files <- paste(path,samples,"gene_exon_tagged_dge.txt.gz",sep = "/")

# Output files:
processed_annotated_file <- "~/Dropbox/Amezcua/processed_data/seurat.merged.processed.rds"
bcells.file <- "~/Dropbox/Amezcua/processed_data/testing.bcells.rds"   # Similar to processed annotated file but containing some scGate bcell maker annotations
panBellFile <-  "~/Dropbox/Amezcua/processed_data/PanBcellSubset.rds"  # Subseted Pan Bcell FIle with recomputed UMAP
```

## Load proccessed object
```{r}
if(file.exists(processed_annotated_file)){
  run =F
  seurat.merged.processed <- readRDS(processed_annotated_file)  
}else{
  run =T
}

```


#create Seurat objects
```{r}
if(run){
  seuratDataList <- list()
  for(i in seq_along(samples)){
    countdata <- read.table(
      file = count_files[i],
      header = T,
      row.names = 1
    )
      
    obj <- Seurat::CreateSeuratObject(counts = countdata,project = samples[i])
    seuratDataList[[i]] <- obj
  }
  ncell <- table(unlist(sapply(seuratDataList,colnames)))
  to.drop <- names(ncell[ncell>1])
  
  # remove 4 barcodes duplicated across samples
  seuratDataList = lapply(seuratDataList,function(x){
    res = subset(x,cells = setdiff(colnames(x),to.drop) )
    return(res)
  })
}
```


# quality filtering
```{r,fig.height= 8,fig.width= 13}
if(run){
  seurat.merged <- Reduce(merge, seuratDataList)
  seurat.merged@meta.data%>%head()
  
  percent.ribo.dv <- PercentageFeatureSet(seurat.merged, pattern = "^Rp[ls]")
  percent.mito.dv <- PercentageFeatureSet(seurat.merged, pattern = "^mt-")
  
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.ribo.dv, col.name = "percent.ribo")
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.mito.dv, col.name = "percent.mito")
Seurat::VlnPlot(seurat.merged,features = c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),group.by = "orig.ident",ncol = 2)
}  
```

```{r}
#splited.by.sample <- SplitObject(seurat.merged, split.by = "orig.ident")
x <- splited.by.sample[[3]]
q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
q
```

# Filter low or high expressed cells 
```{r}
if(run){
  splited.by.sample <- SplitObject(seurat.merged, split.by = "orig.ident")
  
  splited.by.sample.qc <- lapply(splited.by.sample, function(x) {
     #n UMIs
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nCount_RNA>max(q["10%"]) & nCount_RNA<q['99%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nFeature_RNA>max(q['10%']) & nFeature_RNA < q['99%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.ribo>q['1%'] & percent.ribo < min(50,q['99%']) )
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.mito < q['99%'])
     return(x)
  })
  
  seurat.filtered.merged <- Reduce(merge, splited.by.sample.qc)
  dim(seurat.merged)
  dim(seurat.filtered.merged)
}else{
  seurat.filtered.merged <- seurat.merged.processed
}

sapply(c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),function(x){summary(seurat.filtered.merged[[x]])})

```

## Control step: how many cells were dropped by high expression values? (Re: 262 cells)
```{r}
if(run){
  high.expr.dropped <- lapply(splited.by.sample, function(x){
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_ncount <- subset(x, nCount_RNA>q['99%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_nfeature <- subset(x, subset= nFeature_RNA > q['99%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_ribo <- subset(x, subset=percent.ribo > min(50,q['99%']))
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_mito <- subset(x, subset=percent.mito > q['99%'])
     
     return(table(c(x_ncount%>%colnames(), x_nfeature%>%colnames(),x_ribo%>%colnames(),x_ribo%>%colnames())))
    }
  )
  
  # 262 cells dropped by staying over the top 1% quantile in any of control quantity
  cells.dropped.by.hihg.quantile <- unlist(high.expr.dropped)
  cells.dropped.by.hihg.quantile%>%length()
}
```


## Sanity check: visual detection of possible bach effects (Re: thats ok)
```{r,fig.height= 4.5,fig.width= 12}
ndim = 20      # for pca
nfeat = 1000    # number of most variable features to be considered

# scale data; find variable features, run pca and run umap
if(run) seurat.merged.processed <- process_obj(seurat.filtered.merged,ndim = ndim, nfeat = nfeat, find_variable_features = T) 
DimPlot(seurat.merged.processed,,split.by = "orig.ident")
```

# 7980 cells
# 18.886 features 
```{r}
seurat.merged.processed
```

## run singleR
```{r}
if(run){
  ref.se <- celldex::ImmGenData()
  param <- BiocParallel::MulticoreParam(workers = 32)
  
  ## broad 
  result.singler.main = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se, labels = ref.se$label.main, BPPARAM=param)
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler.main$pruned.labels,col.name = "SingleR.main")
  
  # finer (ontology based)
  result.singler = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se,labels = ref.se$label.ont, BPPARAM=param)
  # get term names
  cl <- ontoProc::getCellOnto()
  terms <- result.singler$pruned.labels%>%unique()
  getTerms <- ontoProc::label_TAG(ontology = cl,terms)
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler$pruned.labels,col.name = "SingleR.tag")
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = unname(getTerms[result.singler$pruned.labels]),col.name = "SingleR.term")
}

```

# visualize SingleR classification

```{r,fig.height= 8,fig.width= 8}

fig0 = DimPlot(seurat.merged.processed,group.by = "SingleR.main",label =T,repel =T) + theme(aspect.ratio = 1)
ggsave("plots/fig0_singleR.pdf",plot = fig0,width = 7, height = 7)
ggsave("plots/fig0_singleR.png",plot = fig0,width = 7, height = 7)
ggsave("plots/fig0_singleR.tiff",plot = fig0,width = 7, height = 7)
fig0
#(DimPlot(seurat.merged.processed,group.by = "SingleR.term")) +  theme(legend.text=element_text(size=16)) #+ NoLegend())

```


## Clustering

```{r}
if(run){
  set.seed(1234)
  seurat.merged.processed <- FindNeighbors(seurat.merged.processed, dims = 1:20)
  seurat.merged.processed <- FindClusters(seurat.merged.processed, resolution = 0.7)
  saveRDS(seurat.merged.processed,file = processed_annotated_file)
}
DimPlot(seurat.merged.processed,group.by = "seurat_clusters",label = T) + theme(aspect.ratio = 1)
```

```{r}

if(file.exists(bcells.file)){
  run.bcell.markers <- F
  testing.bcells <- readRDS(bcells.file)
}else{
  run.bcell.markers <- T
  testing.bcells <- seurat.merged.processed
}

```

# Here We define our final gating model (panBcell2 which includes plasmatic resident cells)
```{r, fig.height= 4, fig.width= 5}
panBcell2 <- scGate::gating_model(name = "cd79", signature = "Cd79a")
panBcell2 <- scGate::gating_model(panBcell2,name = "cd19", signature = "Cd19")
panBcell2 <- scGate::gating_model(panBcell2, name = "cd20", signature = "Ms4a1")
panBcell2 <- scGate::gating_model(panBcell2, name = "plasmaticas_residentes", signature = "Sdc1;Igha;Cd3e-;Cd3d-;Cd4-;Cd8a-")
```


```{r}
testing.bcells <- scGate(data = testing.bcells, model = panBcell2, output.col.name = "panBcell2")
#DimPlot(testing.bcells,group.by = "panBcell2")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell2" ) + theme(aspect.ratio = 1) + ggtitle("panBcell2")

```

```{r}
seurat.pan.Bcells <- seurat.merged.processed%>%subset(seurat_clusters %in% c(0,1,10,13,17,18))  # CLUSTER 16, REMOVIDO (Jun 2022)
seurat.pan.Bcells <- process_obj(seurat.pan.Bcells ,ndim = 20, nfeat = 500, find_variable_features = T)
set.seed(1234)
seurat.pan.Bcells <- FindNeighbors(seurat.pan.Bcells, dims = 1:20)
seurat.pan.Bcells$original_seurat_clusters <- seurat.pan.Bcells$seurat_clusters
seurat.pan.Bcells <- FindClusters(seurat.pan.Bcells, resolution = 0.4)
```

```{r, fig.height= 4.5,fig.width= 11}
A <- DimPlot(seurat.pan.Bcells, group.by = "original_seurat_clusters")
seurat.pan.Bcells <- scGate(data = seurat.pan.Bcells, model = panBcell2, output.col.name = "panBcell2HR")
B <- DimPlot(seurat.pan.Bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell2HR" ) + theme(aspect.ratio = 1) + ggtitle("panBcell2HR")
A + B
```

## las panBcell2HR se construcyeron a partir de un modelo de scGate pero sobre un subset del dataset -filtrando los clusters con señal (0,1,10,13,16,17,18) --
## Entonces para plotear sobre el dataset completo las HR, tenemos que matchearlas. Eso es lo que se hace en esta celda siguiente

```{r}
testing.bcells$panBcell2HR <- "Impure"
aux <- rep("Impure",dim(testing.bcells)[2])
names(aux) = testing.bcells@meta.data%>%rownames()
panbcellhr = rownames(seurat.pan.Bcells@meta.data[seurat.pan.Bcells$panBcell2HR=="Pure",])
aux[panbcellhr]<- "Pure"
testing.bcells$panBcell2HR <- aux

```


# Write down original seurat clusters
```{r}
original_seurat_clusters <- testing.bcells@meta.data%>%select("seurat_clusters")
write.table(original_seurat_clusters,file = "~/Dropbox/Amezcua/tissue_resident_cells/reports/original_seurat_cluster.tsv",sep = "\t",col.names = T,row.names =T,quote =F)
```


# Plot Fig1
```{r}
fig1a <- DimPlot(testing.bcells,group.by = "seurat_clusters",label = T) + theme(aspect.ratio = 1)
fig1b <- DimPlot(testing.bcells,cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell2HR" ) + theme(aspect.ratio = 1) + ggtitle("panBcell2HR")

ggsave("plots/fig1a_clustering.pdf",plot = fig1a,width = 4.5, height = 4) 
ggsave("plots/fig1a_clustering.png",plot = fig1a,width = 4.5, height = 4)
ggsave("plots/fig1a_clustering.tiff",plot = fig1a,width = 4.5, height = 4)

ggsave("plots/fig1b_PanBcell.pdf",plot = fig1b,width = 4.5, height = 4) 
ggsave("plots/fig1b_PanBcell.png",plot = fig1b,width = 4.5, height = 4)
ggsave("plots/fig1b_PanBcell.tiff",plot = fig1b,width = 4.5, height = 4)

fig1b
```
