# This notebook will work on data produced by Amezcua et al. (Cell 2019)
In that work, Amezcua analyzed tissue resident Th cells. 
Now, the idea is to recycle those data but redirecting the focus on Bcells. In particular, those tissue resident ones. 


```{r}
renv::activate()
renv::restore()
#install.packages("Seurat")
#install.packages("remotes")
#remotes::install_github("carmonalab/UCell")
#remotes::install_github("carmonalab/scGate")
#install.packages("BiocManager")
#BiocManager::install("SingleR")
#BiocManager::install("celldex")
#BiocManager::install("ontoProc")
#BiocManager::install("rmarkdown")

library("Seurat")
library("dplyr")
library("UCell")
library("scGate")
library("SingleR")
library("ggplot2")
source("./utils.R")

```

# Load datasets from counts
```{r}
path <- c("~/Dropbox/Amezcua/data/")
samples <- c("DS-017-WT","DS-017-Im1","DS-017-Im2")
count_files <- paste(path,samples,"gene_exon_tagged_dge.txt.gz",sep = "/")

# Output files:
processed_annotated_file <- "~/Dropbox/Amezcua/processed_data/seurat.merged.processed.rds"
bcells.file <- "../processed_data/testing.bcells.rds"   # Similar to processed annotated file but containing some scGate bcell maker annotations
panBellFile <-  "../processed_data/PanBcellSubset.rds"  # Subseted Pan Bcell FIle with recomputed UMAP
```

## Load proccessed object
```{r}
if(file.exists(processed_annotated_file)){
  run =F
  seurat.merged.processed <- readRDS(processed_annotated_file)  
}else{
  run =T
}

```


#create Seurat objects
```{r}
if(run){
  seuratDataList <- list()
  for(i in seq_along(samples)){
    countdata <- read.table(
      file = count_files[i],
      header = T,
      row.names = 1
    )
      
    obj <- Seurat::CreateSeuratObject(counts = countdata,project = samples[i])
    seuratDataList[[i]] <- obj
  }
  ncell <- table(unlist(sapply(seuratDataList,colnames)))
  to.drop <- names(ncell[ncell>1])
  
  # remove 4 barcodes duplicated across samples
  seuratDataList = lapply(seuratDataList,function(x){
    res = subset(x,cells = setdiff(colnames(x),to.drop) )
    return(res)
  })
}
```


# quality filtering
```{r,fig.height= 8,fig.width= 13}
if(run){
  seurat.merged <- Reduce(merge, seuratDataList)
  seurat.merged@meta.data%>%head()
  
  percent.ribo.dv <- PercentageFeatureSet(seurat.merged, pattern = "^Rp[ls]")
  percent.mito.dv <- PercentageFeatureSet(seurat.merged, pattern = "^mt-")
  
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.ribo.dv, col.name = "percent.ribo")
  seurat.merged <- AddMetaData(seurat.merged, metadata = percent.mito.dv, col.name = "percent.mito")
}  
Seurat::VlnPlot(seurat.merged,features = c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),group.by = "orig.ident",ncol = 2)
```

# Filter low or high expressed cells 
```{r}
if(run){
  splited.by.sample <- SplitObject(seurat.merged, split.by = "orig.ident")
  
  splited.by.sample.qc <- lapply(splited.by.sample, function(x) {
     #n UMIs
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nCount_RNA>max(q["10%"]) & nCount_RNA<q['99%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99))
     x <- subset(x, subset=nFeature_RNA>max(q['10%']) & nFeature_RNA < q['99%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.ribo>q['1%'] & percent.ribo < min(50,q['99%']) )
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.9,0.95,0.98,0.99),na.rm =T)
     x <- subset(x, subset=percent.mito < q['99%'])
     return(x)
  })
  
  seurat.filtered.merged <- Reduce(merge, splited.by.sample.qc)
  dim(seurat.merged)
  dim(seurat.filtered.merged)
}else{
  seurat.filtered.merged <- seurat.merged.processed
}

sapply(c("nCount_RNA","nFeature_RNA","percent.mito", "percent.ribo"),function(x){summary(seurat.filtered.merged[[x]])})

```

## Control step: how many cells were dropped by high expression values? (Re: 262 cells)
```{r}
if(run){
  high.expr.dropped <- lapply(splited.by.sample, function(x){
     q <- quantile(x@meta.data$nCount_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_ncount <- subset(x, nCount_RNA>q['99%']) 
     
     #n genes
     q <- quantile(x@meta.data$nFeature_RNA, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99))
     x_nfeature <- subset(x, subset= nFeature_RNA > q['99%'])
     
     #% ribo
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_ribo <- subset(x, subset=percent.ribo > min(50,q['99%']))
     
     #% mito
     q <- quantile(x@meta.data$percent.ribo, c(0.01,0.02,0.05,0.1,0.5,0.9,0.95,0.98,0.99),na.rm =T)
     x_mito <- subset(x, subset=percent.mito > q['99%'])
     
     return(table(c(x_ncount%>%colnames(), x_nfeature%>%colnames(),x_ribo%>%colnames(),x_ribo%>%colnames())))
    }
  )
  
  # 262 cells dropped by staying over the top 1% quantile in any of control quantity
  cells.dropped.by.hihg.quantile <- unlist(high.expr.dropped)
  cells.dropped.by.hihg.quantile%>%length()
}
```


## Sanity check: visual detection of possible bach effects (Re: thats ok)
```{r,fig.height= 4.5,fig.width= 12}
ndim = 20      # for pca
nfeat = 1000    # number of most variable features to be considered

# scale data; find variable features, run pca and run umap

if(run) seurat.merged.processed <- process_obj(seurat.filtered.merged,ndim = ndim, nfeat = nfeat, find_variable_features = T) 
DimPlot(seurat.merged.processed,,split.by = "orig.ident")
```

# 7980 cells
# 18.886 features 
```{r}
seurat.merged.processed
```

## run singleR
```{r}
if(run){
  ref.se <- celldex::ImmGenData()
  param <- BiocParallel::MulticoreParam(workers = 32)
  
  ## broad 
  result.singler.main = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se, labels = ref.se$label.main, BPPARAM=param)
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler.main$pruned.labels,col.name = "SingleR.main")
  
  # finer (ontology based)
  result.singler = SingleR(test = GetAssayData(seurat.merged.processed), ref = ref.se,labels = ref.se$label.ont, BPPARAM=param)
  # get term names
  cl <- ontoProc::getCellOnto()
  terms <- result.singler$pruned.labels%>%unique()
  getTerms <- ontoProc::label_TAG(ontology = cl,terms)
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = result.singler$pruned.labels,col.name = "SingleR.tag")
  seurat.merged.processed <- AddMetaData(seurat.merged.processed, metadata = unname(getTerms[result.singler$pruned.labels]),col.name = "SingleR.term")
}

```

# visualize SingleR classification

## TO DO, reahacer cluster por ontologia sólo en Bcells y Plasma Cells. 
```{r,fig.height= 8,fig.width= 30}
#DimPlot(seurat.merged.processed,group.by = "SingleR.main") +  

(DimPlot(seurat.merged.processed,group.by = "SingleR.term")) +  theme(legend.text=element_text(size=16)) #+ NoLegend())

```


## explore some key markers
```{r}
models <- scGate::get_scGateDB()
scg.immune <- scGate::gating_model(name = "immune", signature = "Ptprc")
panBcell <- scGate::gating_model(name = "panBcell", signature = "Cd79a")
```

Run
```{r}
if(run){
  seurat.merged.processed <- scGate(data = seurat.merged.processed, model = scg.immune, output.col.name = "immune")
  seurat.merged.processed <- scGate(data = seurat.merged.processed, model = models$mouse$generic$Myeloid, output.col.name = "myeloid")
  seurat.merged.processed <- scGate(data = seurat.merged.processed, model = models$mouse$generic$Tcell, output.col.name = "Tcell")
  seurat.merged.processed <- scGate(data = seurat.merged.processed, model = panBcell, output.col.name = "PanBcell")
}
```


```{r}
DimPlot(seurat.merged.processed, cols = c(list(Impure = "gray", Pure = "green")), group.by = "immune" ) + theme(aspect.ratio = 1) + ggtitle("scGate Immune")
DimPlot(seurat.merged.processed, cols = c(list(Impure = "gray", Pure = "green")), group.by = "myeloid" ) + theme(aspect.ratio = 1) + ggtitle("scGate myeloid")
DimPlot(seurat.merged.processed, cols = c(list(Impure = "gray", Pure = "green")), group.by = "Tcell" ) + theme(aspect.ratio = 1) + ggtitle("scGate Tcell")

DimPlot(seurat.merged.processed, cols = c(list(Impure = "gray", Pure = "green")), group.by = "PanBcell" ) + theme(aspect.ratio = 1) + ggtitle("scGate panBcell")

```

## Clustering

```{r}
if(run){
  set.seed(1234)
  seurat.merged.processed <- FindNeighbors(seurat.merged.processed, dims = 1:20)
  seurat.merged.processed <- FindClusters(seurat.merged.processed, resolution = 0.7)
  saveRDS(seurat.merged.processed,file = processed_annotated_file)
}
```

```{r}
DimPlot(seurat.merged.processed,group.by = "seurat_clusters",label = T)

```


# Marker enrichment of clusters
```{r}
Idents(seurat.merged.processed) <- "seurat_clusters"
find.all.markers <- FindAllMarkers(object = seurat.merged.processed, min.diff.pct = 0.5) #0.4
to.review <- find.all.markers%>%subset(cluster %in% c(0,1,10,13,16,17,18))
```

```{r}
aux <- find.all.markers%>%subset(cluster == 13)
aux%>%rownames() %>%grep("S",., value = T)
```

# see immunoglobulines
```{r,fig.height= 5,fig.width= 6}
ighs <- data.frame() 
ighs <- rbind( ighs, c("Igm","Ighm"))
colnames(ighs) <- c("protein","gene")
ighs <- rbind(ighs, c("Iga","Igha"))
ighs <- rbind(ighs, c("Igd","Ighd1-1"))
ighs <- rbind(ighs, c("Igg","Ighg1"))
FeaturePlot(seurat.merged.processed,features = ighs$gene,pt.size = 0.25,ncol = 2)



```




```{r,fig.height= 8,fig.width= 12}
markers <- data.frame()
markers <- rbind(markers,c("Cd138","Sdc1"))
colnames(markers) <- c("protein","gene")
markers <- rbind(markers, c("Cd79a","Cd79a"))
markers <- rbind(markers, c("Cd20","Ms4a1"))
markers <- rbind(markers, c("Cd19","Cd19"))
markers <- rbind(markers, c("Cd23","Fcer2a"))
markers <- rbind(markers, c("Cd93","Cd93"))



markers$gene %in% (seurat.merged.processed%>%rownames())
#seurat.merged.processed%>%rownames()%>%grep('Cd8',.,value = T)
FeaturePlot(seurat.merged.processed,features = markers$gene,pt.size = 0.5,ncol = 3) 

```
```{r}
cd79 <- scGate::gating_model(name = "cd79", signature = "Cd79a")
cd19 <- scGate::gating_model(name = "cd19", signature = "Cd19")
cd20 <- scGate::gating_model(name = "cd20", signature = "Ms4a1")
cd23 <- scGate::gating_model(name = "cd23", signature = "Fcer2a")
cd93 <- scGate::gating_model(name = "cd93", signature = "Cd93")
cd138 <- scGate::gating_model(name = "cd138", signature = "Sdc1")
cd80 <- scGate::gating_model(name = "cd80", signature = "Cd80")
cd86 <- scGate::gating_model(name = "cd86", signature = "Cd86")
igg <- scGate::gating_model(name = "igg", signature = "Ighg1")
igm <- scGate::gating_model(name = "igm", signature = "Ighm")
iga <- scGate::gating_model(name = "iga", signature = "Igha")
igd <- scGate::gating_model(name = "igd", signature = "Ighd1-1")
```


```{r}
if(file.exists(bcells.file)){
  run.bcell.markers <- F
  testing.bcells <- readRDS(bcells.file)
}else{
  run.bcell.markers <- T
  testing.bcells <- seurat.merged.processed

}

```

```{r}
## new scGate model
cd79 <- scGate::gating_model(name = "cd79", signature = "Cd79a")
cd19 <- scGate::gating_model(name = "cd19", signature = "Cd19")
cd20 <- scGate::gating_model(name = "cd20", signature = "Ms4a1")
cd23 <- scGate::gating_model(name = "cd23", signature = "Fcer2a")
cd93 <- scGate::gating_model(name = "cd93", signature = "Cd93")
cd138 <- scGate::gating_model(name = "cd138", signature = "Sdc1")
cd80 <- scGate::gating_model(name = "cd80", signature = "Cd80")
cd86 <- scGate::gating_model(name = "cd86", signature = "Cd86")
igg <- scGate::gating_model(name = "igg", signature = "Ighg1")
igm <- scGate::gating_model(name = "igm", signature = "Ighm")
iga <- scGate::gating_model(name = "iga", signature = "Igha")
igd <- scGate::gating_model(name = "igd", signature = "Ighd1-1")


if(run.bcell.markers){
  testing.bcells <- scGate(data = testing.bcells, model = cd79, output.col.name = "cd79")
  testing.bcells <- scGate(data = testing.bcells, model = cd19, output.col.name = "cd19")
  testing.bcells <- scGate(data = testing.bcells, model = cd20, output.col.name = "cd20")
  testing.bcells <- scGate(data = testing.bcells, model = cd23, output.col.name = "Fcer2a")
  testing.bcells <- scGate(data = testing.bcells, model = cd93, output.col.name = "cd93")
  testing.bcells <- scGate(data = testing.bcells, model = cd138, output.col.name = "cd138")
  testing.bcells <- scGate(data = testing.bcells, model = cd80, output.col.name = "cd80")
  testing.bcells <- scGate(data = testing.bcells, model = cd86, output.col.name = "cd86")
  testing.bcells <- scGate(data = testing.bcells, model = igm, output.col.name = "igm")
  testing.bcells <- scGate(data = testing.bcells, model = igg, output.col.name = "igg")
  testing.bcells <- scGate(data = testing.bcells, model = iga, output.col.name = "iga")
  testing.bcells <- scGate(data = testing.bcells, model = igd, output.col.name = "igd")
}


DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd79" ) + theme(aspect.ratio = 1) + ggtitle("cd79")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd19" ) + theme(aspect.ratio = 1) + ggtitle("cd19")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd20" ) + theme(aspect.ratio = 1) + ggtitle("cd20")
#DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd23" ) + theme(aspect.ratio = 1) + ggtitle("cd23")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd93" ) + theme(aspect.ratio = 1) + ggtitle("cd93")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd138" ) + theme(aspect.ratio = 1) + ggtitle("cd138")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd80" ) + theme(aspect.ratio = 1) + ggtitle("cd80")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "cd86" ) + theme(aspect.ratio = 1) + ggtitle("cd86")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "igm" ) + theme(aspect.ratio = 1) + ggtitle("igm")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "igg" ) + theme(aspect.ratio = 1) + ggtitle("igg")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "iga" ) + theme(aspect.ratio = 1) + ggtitle("iga")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "igd" ) + theme(aspect.ratio = 1) + ggtitle("igd")



```

```{r, fig.height= 4, fig.width= 5}
sum(testing.bcells$cd79 == "Pure" | testing.bcells$cd20 == "Pure" | testing.bcells$cd19== "Pure")
sum(testing.bcells$cd79 == "Pure" | testing.bcells$cd20 == "Pure" )


panBcell2 <- scGate::gating_model(name = "cd79", signature = "Cd79a")
panBcell2 <- scGate::gating_model(panBcell2,name = "cd19", signature = "Cd19")
panBcell2 <- scGate::gating_model(panBcell2, name = "cd20", signature = "Ms4a1")
panBcell2 <- scGate::gating_model(panBcell2, name = "plasmaticas_residentes", signature = "Sdc1;Igha;Cd3e-;Cd3d-;Cd4-;Cd8a-")
```

# plasmaticas residentes
```{r}
plasm.resid <- scGate::gating_model( name = "plasmaticas_residentes", signature = "Sdc1;Igha;Cd3e-;Cd3d-;Cd4-;Cd8a-")
if(run.bcell.markers){
  testing.bcells <- scGate(data = testing.bcells, model = plasm.resid, output.col.name = "plasm.resid")
}
```

```{r, fig.height= 4, fig.width= 5}
if(run.bcell.markers) {
  testing.bcells <- scGate(data = testing.bcells, model = panBcell2, output.col.name = "panBcell2")
  saveRDS(testing.bcells,file = "../processed_data/testing.bcells.rds")
}
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell2" ) + theme(aspect.ratio = 1) + ggtitle("panBcell2")
DimPlot(testing.bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "plasm.resid" ) + theme(aspect.ratio = 1) + ggtitle("plasm.resid")

```

* Notas charlando con Caro: prestarle atención a esto
Igj; IGJ   Union de dímero de IGA y pentámero de IGM
Tnfrsf17 : BCMA  receptor expresado en PlasmaCell, señal antiapotótica


## Subset Bcells and plot with the same reduction
```{r, fig.height= 4.5,fig.width= 11}
seurat.pan.Bcells.raw <- seurat.merged.processed%>%subset(seurat_clusters %in% c(0,1,10,13,16,17,18))
Araw <- DimPlot(seurat.pan.Bcells.raw, group.by = "seurat_clusters")
seurat.pan.Bcells.raw <- scGate(data = seurat.pan.Bcells.raw, model = panBcell2, output.col.name = "panBcell2HR")
Braw <- DimPlot(seurat.pan.Bcells.raw, cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell2HR" ) + theme(aspect.ratio = 1) + ggtitle("panBcell2HR")
Araw + Braw
```

```{r}
seurat.pan.Bcells <- seurat.merged.processed%>%subset(seurat_clusters %in% c(0,1,10,13,16,17,18))
seurat.pan.Bcells <- process_obj(seurat.pan.Bcells ,ndim = 20, nfeat = 500, find_variable_features = T)
set.seed(1234)
seurat.pan.Bcells <- FindNeighbors(seurat.pan.Bcells, dims = 1:20)
seurat.pan.Bcells <- FindClusters(seurat.pan.Bcells, resolution = 0.4)
```

```{r, fig.height= 4.5,fig.width= 11}
A <- DimPlot(seurat.pan.Bcells, group.by = "seurat_clusters")
seurat.pan.Bcells <- scGate(data = seurat.pan.Bcells, model = panBcell2, output.col.name = "panBcell2HR")
B <- DimPlot(seurat.pan.Bcells, cols = c(list(Impure = "gray", Pure = "green")), group.by = "panBcell2HR" ) + theme(aspect.ratio = 1) + ggtitle("panBcell2HR")
A + B
```


## Save Pan Bcell data    
```{r}
saveRDS(seurat.pan.Bcells,file = panBcellFile)
```

# I Tried to run SingleR on this dataset but results only make a bit sence at SingleR.main resolution level (i.e. : no use)
```{r}
if(F){
  ref.se <- celldex::ImmGenData()
  param <- BiocParallel::MulticoreParam(workers = 32)
  
  ## broad 
  result.singler.main = SingleR(test = GetAssayData(seurat.pan.Bcells), ref = ref.se, labels = ref.se$label.main, BPPARAM=param)
  seurat.pan.Bcells <- AddMetaData(seurat.pan.Bcells, metadata = result.singler.main$pruned.labels,col.name = "SingleR.main")
  
  # finer (ontology based)
  result.singler = SingleR(test = GetAssayData(seurat.pan.Bcells), ref = ref.se,labels = ref.se$label.ont, BPPARAM=param)
  # get term names
  cl <- ontoProc::getCellOnto()
  terms <- result.singler$pruned.labels%>%unique()
  getTerms <- ontoProc::label_TAG(ontology = cl,terms)
  seurat.pan.Bcells <- AddMetaData(seurat.pan.Bcells, metadata = result.singler$pruned.labels,col.name = "SingleR.tag")
  seurat.pan.Bcells <- AddMetaData(seurat.pan.Bcells, metadata = unname(getTerms[result.singler$pruned.labels]),col.name = "SingleR.term")

seurat.pan.Bcells$SingleR.term%>%unique
(DimPlot(seurat.pan.Bcells,group.by = "SingleR.main")) +  theme(legend.text=element_text(size=5)) #+ NoLegend())
}

```

```{r, fig.height= 4.5,fig.width= 11}


```

