---
title: "R Notebook"
output: html_notebook
---


```{r}
renv::activate()
renv::restore()

library(Seurat)
library(dplyr)
library(UCell)
library(scGate)
library(ggplot2)
library(patchwork)
library(SeuratDisk)
library(ProjecTILs)
source("./utils.R")
```
# TO DO ##########
# Prefiltrar células acorde a SCGATE PURE
# CALCULAR 20 O 50 TOP VARIABLE FEATURES EN BCELLS DEL DATASET MOUSE ORIGINAL (Y TOMAR LOS QUE TAMBIÉN ESTÉN EN Human)
# Luego usarlos para calcular los anchors de integración. 
# VER SI DA MÁS CONSISTENTE EL ANÁLISIS. 


```{r}
panBellFile <-  "~/Dropbox/Amezcua/processed_data/PanBcellSubset.rds"  # Subseted Pan Bcell FIle with recomputed UMAP
data <- readRDS(panBellFile)
data%>%dim
data <- data%>%subset(panBcell2HR == "Pure")
data%>%dim
```

# Load Refernce
```{r}
download = F
if(download){
  options(timeout=3600)
  download.file("https://atlas.fredhutch.org/data/nygc/multimodal/pbmc_multimodal.h5seurat",destfile = "../data/pbmc_multimodal.h5seurat")
}else{
  reference <- LoadH5Seurat("~/Dropbox/Amezcua/data/pbmc_multimodal.h5seurat")
}

```


```{r}
DimPlot(object = reference, reduction = "wnn.umap", group.by = "celltype.l2", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
```

# mapping mouse to human genes
```{r}
## Mapp via Ensemble
if(F){
  human.mapped.genes <- convertMouseGeneList(data%>%rownames)
  
  # Filter those genes that do not appear in our reference
  human.mapped.genes <- human.mapped.genes%>%subset(HGNC.symbol%in%(reference%>%rownames())) 
  
  # make HGNC.symbol unique
  human.mapped.genes <- human.mapped.genes%>%distinct(.,HGNC.symbol,.keep_all = T) 
  
  # make MGI.symbol unique
  human.mapped.genes <- human.mapped.genes%>%distinct(.,MGI.symbol,.keep_all = T) 
}else{
  translate <- ProjecTILs::Hs2Mm.convert.table
  mouse.genes <- data%>%rownames() 
  ii <- match(mouse.genes,translate$Gene.MM)
  human.mapped.genes <- translate[ii[!is.na(ii)],c("Gene.HS","Gene.MM")]
  colnames(human.mapped.genes) <- c("HGNC.symbol","MGI.symbol")
  
}

data.filtered <- data%>%subset(features =  human.mapped.genes$MGI.symbol)
renamed.counts <- data.filtered@assays$RNA@counts
ii <- match(renamed.counts%>%rownames(),human.mapped.genes$MGI.symbol)
rownames(renamed.counts) <- human.mapped.genes[ii,]$HGNC.symbol 

# create new seurat object 
data.human.genes <- CreateSeuratObject(counts = renamed.counts,meta.data = data.filtered@meta.data)

```



```{r}
data.human.genes <- SCTransform(data.human.genes, verbose = FALSE)
#DefaultAssay(data.human.genes) <- "RNA"
#data.human.genes <- process_obj(obj = data.human.genes, ndim = 20, nfeat = 500, find_variable_features = T)
DefaultAssay(data.human.genes) <- "SCT"
```



```{r}
obj <- data.human.genes
obj <- NormalizeData(obj, verbose = TRUE)
obj <- FindVariableFeatures(obj, selection.method = "vst", nfeatures = 50 , verbose = FALSE)
HVGs_query <- obj@assays$SCT@var.features
HVGs_query

```

#Limit reference
```{r,fig.width= 9, fig.height=4 }
ref <- reference%>%subset(celltype.l1 == "B")
DimPlot(object = ref, reduction = "wnn.umap", group.by = "celltype.l2", label = TRUE, label.size = 3, repel = TRUE) + NoLegend() + 
  
DimPlot(object = ref, reduction = "wnn.umap", group.by = "celltype.l3", label = TRUE, label.size = 3, repel = TRUE) 


```

```{r}
obj.ref <- FindVariableFeatures(ref, selection.method = "vst", nfeatures = 5000 , verbose = FALSE,assay = "SCT")
HVGs_ref <- obj.ref@assays$SCT@var.features
HVGs_ref%>%head(20)
```


```{r}
intersect(HVGs_ref,HVGs_query)
intersect(HVGs_ref,data.human.genes%>%rownames)%>%length()
```

```{r}
length(intersect(reference@assays$SCT@var.features, HVGs_ref))
```


Find Anchors
```{r}
anchors <- FindTransferAnchors(
  reference = ref,
  query = data.human.genes,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50,
  features = intersect(HVGs_ref,data.human.genes%>%rownames)
)
```


Mapping
```{r}
mapped.data <- MapQuery(
  anchorset = anchors,
  query = data.human.genes,
  reference = ref,
  refdata = list(
    celltype.l1 = "celltype.l1",
    celltype.l2 = "celltype.l2",
    celltype.l3 = "celltype.l3",
    predicted_ADT = "ADT"
  ),
  reference.reduction = "spca", 
  reduction.model = "wnn.umap"
)
```

# Explore results
```{r,fig.width= 15,fig.height=5}
p1 = DimPlot(mapped.data, reduction = "ref.umap", group.by = "predicted.celltype.l1", label = TRUE, label.size = 3, repel = TRUE) + NoLegend()
p2 = DimPlot(mapped.data, reduction = "ref.umap", group.by = "predicted.celltype.l2", label = TRUE, label.size = 3 ,repel = TRUE) + NoLegend()
p3 = DimPlot(mapped.data, reduction = "ref.umap", group.by = "predicted.celltype.l3", label = TRUE, label.size = 3 ,repel = TRUE) 

p1 + p2 +p3
```


See some marker genes
```{r,fig.width= 10,fig.height=4}
tt <- table(mapped.data@meta.data$predicted.celltype.l2,mapped.data@meta.data$panBcell2HR)
tt
```


```{r,fig.width= 10,fig.height=4}
FeaturePlot(mapped.data, features = c("B naive", "B intermediate", "B memory"),  reduction = "ref.umap", cols = c("lightgrey", "darkred"), ncol = 3) & theme(plot.title = element_text(size = 10))

FeaturePlot(pures, features = c("B naive", "B intermediate", "B memory"),  reduction = "ref.umap", cols = c("lightgrey", "darkred"), ncol = 3) & theme(plot.title = element_text(size = 10))

```


## Define a final working dataset
```{r}
low_res = c("B")
medium_res = c("B naive", "B intermediate", "B memory", "Plasmablast")
high_res <- c("B intermediate kappa", "B intermediate lambda","B memory kappa","B memory lambda","B naive kappa", "B naive lambda", "Plasma","Plasmablast")

# Filter
mapped.data.filtered <- mapped.data %>% subset(.,predicted.celltype.l1 %in% low_res &
                                                 predicted.celltype.l2 %in% medium_res &
                                                 predicted.celltype.l3 %in% high_res & 
                                                 panBcell2HR =="Pure"
)
mapped.data%>%dim
mapped.data.filtered%>%dim

```


# Visualize filtered data
```{r,fig.width=8,fig.height=4}
p4 <- DimPlot(mapped.data.filtered, reduction = "ref.umap", group.by = "predicted.celltype.l3", label = TRUE, label.size = 3 ,repel = TRUE) + NoLegend()
p5 <- FeaturePlot(mapped.data.filtered, reduction = "ref.umap", features = "predicted.celltype.l1.score", label = TRUE, label.size = 3 ,repel = TRUE) 
p4 + p5
```


```{r}
mapped.data.filtered$predicted.celltype.l1.score%>%hist()
```


```{r}
quantile(mapped.data.filtered$predicted.celltype.l1.score,c(0.01,0.03,0.05))
q01 <- quantile(mapped.data.filtered$predicted.celltype.l1.score,0.01)
q03 <- quantile(mapped.data.filtered$predicted.celltype.l1.score,0.03)
q05 <- quantile(mapped.data.filtered$predicted.celltype.l1.score,0.05)

mapped.data.filtered<- mapped.data.filtered%>%subset(predicted.celltype.l1.score >q05)
p6 <-DimPlot(mapped.data.filtered, reduction = "ref.umap", group.by  = "predicted.celltype.l3", label = TRUE, label.size = 3 ,repel = TRUE) 
p6
```

How many cells mapped to each state?
```{r}
mapped.data.filtered$predicted.celltype.l3%>%table
```

# Transfer learned celltypes to the mouse original data
```{r}
cells.ok <- colnames(data)%in% colnames(mapped.data.filtered)
mouse.data.mapped <- data%>%subset(cells = intersect(colnames(data),colnames(mapped.data.filtered)))

mouse.data.mapped <- AddMetaData(mouse.data.mapped, metadata = mapped.data.filtered@meta.data$predicted.celltype.l1, col.name = "predicted.celltype.l1")

mouse.data.mapped <- AddMetaData(mouse.data.mapped, metadata = mapped.data.filtered@meta.data$predicted.celltype.l1.score, col.name = "predicted.celltype.l1.score")


mouse.data.mapped <- AddMetaData(mouse.data.mapped, metadata = mapped.data.filtered@meta.data$predicted.celltype.l2, col.name = "predicted.celltype.l2")
mouse.data.mapped <- AddMetaData(mouse.data.mapped, metadata = mapped.data.filtered@meta.data$predicted.celltype.l2.score, col.name = "predicted.celltype.l2.score")


mouse.data.mapped <- AddMetaData(mouse.data.mapped, metadata = mapped.data.filtered@meta.data$predicted.celltype.l3, col.name = "predicted.celltype.l3")
mouse.data.mapped <- AddMetaData(mouse.data.mapped, metadata = mapped.data.filtered@meta.data$predicted.celltype.l3.score, col.name = "predicted.celltype.l3.score")

```


## We can explore some markers among defined celltypes

```{r}
#mouse.data.mapped%>%dim
mouse.data.mapped%>%rownames()%>%grep("^Igl",.,value = T)

```

```{r}

VlnPlot(mouse.data.mapped, features = c("Ighm", "Ighg1", "Igha","Cd19","Ms4a1","Cd38","Cd69","Cd93","Sdc1","Itgal","Igkc","Iglc2"), group.by = "predicted.celltype.l3", stack = T) + NoLegend()


mouse.data.mapped$predicted.celltype.l3%>%table(exclude = F)
```
```{r}
p7 <-DimPlot(mouse.data.mapped, reduction = "umap", group.by  = "predicted.celltype.l3", label = TRUE, label.size = 3 ,repel = TRUE) 
p7
```

