---
title: "Exploring Bcell markers in Amezcua 2019 dataset"
author: "A. Berenstein"
date: "22/04/2022"
output:
  rmdformats::readthedown:
    self-contained: true
    highlight: haddock
    thumbnails: false
    css: styles.css
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file, encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'explore_cluster_signatures.html'))})
---


```{r, message=F, warning=F,results=F}
renv::activate()
renv::restore()

library(Seurat)
library(dplyr)
library(UCell)
library(scGate)
library(ggplot2)
library(patchwork)
library(SeuratDisk)
library(ProjecTILs)
library(glmGamPoi)
source("./utils.R")
```



# Load Bcell-filtered dataset
```{r,warning=F, collapse =T,message =F}
ndim = 30; seed = 31416
panBellFile <-  "~/Dropbox/Amezcua/processed_data/PanBcellSubset.rds"  # Subseted Pan Bcell FIle with recomputed UMAP
data <- readRDS(panBellFile)
data%>%dim
data <- data%>%subset(panBcell2HR == "Pure")
data%>%dim
data$condition <- ifelse(data$orig.ident == "DS-017-WT","wt","im") 
data.bkp <- data
data.wt <- data%>%subset(orig.ident == "DS-017-WT")
data.im <- data%>%subset(orig.ident != "DS-017-WT")
```

Visualize datasets by condition
```{r,fig.width=6,fig.height=3,warning=F, collapse =T,message =F}
DimPlot(data,group.by = "orig.ident",split.by = "orig.ident")
```

# Preprocessing

**We will preprocess separately WT and IM subsets**

## Cell cycle scoring and regression 

The idea here is to mitigate the effects of cell cycle heterogeneity by calculating cell cycle phase scores based on canonical markers, and regressing these out of the data during pre-processing (see [SeuratVignette cell cycle](https://satijalab.org/seurat/articles/cell_cycle_vignette.html) for details)

```{r,warning=F, collapse =T,message =F, results='hide'}
# extract cell cycle genes from Seurat 
s.genes <- cc.genes$s.genes%>%hs.2.mm(.)
g2m.genes <- cc.genes$g2m.genes%>%hs.2.mm(.)
```


```{r,warning=F, collapse =T,message =F,results='hide'}
# IM
set.seed(seed)
data.im <- CellCycleScoring(data.im, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.im$CC.Difference <- data.im$S.Score - data.im$G2M.Score
data.im <- SCTransform(data.im, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE)
data.im <- RunPCA(data.im, features = data.im@assays$RNA@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.im <- RunUMAP(data.im, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)

```


```{r,warning=F, collapse =T,message =F, results='hide'}
# WT
set.seed(seed)
data.wt <- CellCycleScoring(data.wt, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.wt$CC.Difference <- data.wt$S.Score - data.wt$G2M.Score
data.wt <- SCTransform(data.wt, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE)
data.wt <- RunPCA(data.wt, features = data.wt@assays$RNA@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.wt <- RunUMAP(data.wt, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)


```


# Unsupervised analysis
```{r,fig.height=4, fig.width= 10,warning=F, collapse =T,message =F, results='hide'}
set.seed(seed)
data.im <- FindNeighbors(data.im,assay  = "SCT", dims = 1:20)
data.im <- FindClusters(data.im, resolution = 0.7)
data.im <- FindClusters(data.im, resolution = 0.5)
data.im <- FindClusters(data.im, resolution = 0.4)
data.im <- FindClusters(data.im, resolution = 0.3)

set.seed(seed)
data.wt <- FindNeighbors(data.wt,assay  = "SCT", dims = 1:20)
data.wt <- FindClusters(data.wt, resolution = 0.7)
data.wt <- FindClusters(data.wt, resolution = 0.5)
data.wt <- FindClusters(data.wt, resolution = 0.4)
data.wt <- FindClusters(data.wt, resolution = 0.3)

a <- DimPlot(data.wt,group.by = "SCT_snn_res.0.3",label = T)
b <- DimPlot(data.im ,group.by = "SCT_snn_res.0.3",label = T)
a + b
```


```{r,fig.height= 12,fig.width=10,warning=F, collapse =T,message =F}
Idents(data.wt) <- "SCT_snn_res.0.3"
discrim.wt <- FindAllMarkers(data.wt, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
gene.wt <- discrim.wt$gene
d <- DoHeatmap(data.wt,features = gene.wt)

Idents(data.im) <- "SCT_snn_res.0.3"
discrim.im <- FindAllMarkers(data.im,assay = "SCT", only.pos = T, min.pct = 0.40, logfc.threshold = 0.25)
gene.im <- discrim.im$gene
e <- DoHeatmap(data.im,features = gene.im,size = 3,lines.width = 35)

# show WT heatmap
d 

# show IM heatmap
e 
 
```


# DimPlot using Heatmap genes

Plot again clusters in WT and IM datasets
```{r,fig.height=4, fig.width= 10,warning=F, collapse =T,message =F, results='hide'}
a + b
```

WT dataset
```{r,collapse =T,fig.width=14,fig.height=35,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = discrim.wt$gene,ncol = 5)
```

IM dataset
```{r,fig.width= 15,fig.height= 40,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = intersect(discrim.im$gene,data.im%>%rownames()),ncol = 5)
```


# Explore predefined markers 

We will use here those genes previously defined in our molecular signature spreadsheet file
```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
features = c("Ighm","Ighd1-1","Igha","Ighg1","Cd19","Ms4a1","Fcer2a","Cd24a","Cd38","Pdcd1lg2","Nt5e","Entpd1","Cd80","Cd86","Cd93","Cd69","Itgal","Cd44","Cd28","Sdc1",	"Prdm1","Tnfrsf17","Sell","Igj")
```

WT dataset
```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = features, ncol = 5)
```

IM dataset
```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = features, ncol = 5)
```


# Explore markers of resident plasmatic cells (IgA+)
WT dataset
```{r,fig.width= 12,fig.height=8,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```

IM dataset
```{r,fig.width= 13,fig.height=7,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```

Include scGate model
```{r,fig.height=4,fig.width=8,warning=F, collapse =T,message =F}
plasm.resid.4 <- scGate::gating_model(name = "plasmaticas_residentes", signature = "Sdc1;Igha;Prdm1;Tnfrsf17;Igj")
data.im <- scGate(data = data.im, model = plasm.resid.4, output.col.name = "plasm.resid.4")
p4.im <- DimPlot(data.im, cols = c(list(Impure = "gray", Pure = "green")), group.by = "plasm.resid.4") + theme(aspect.ratio = 1) + ggtitle("plasma resid") + 
  labs(subtitle =  "Cd138+ Igha+ Blimp-1+ BCMA+ Igj+",caption = sprintf("Ncells: %s",sum(data.im$plasm.resid.4=="Pure")))

data.wt <- scGate(data = data.wt, model = plasm.resid.4, output.col.name = "plasm.resid.4")
p4.wt <- DimPlot(data.wt, cols = c(list(Impure = "gray", Pure = "green")), group.by = "plasm.resid.4") + theme(aspect.ratio = 1) + ggtitle("plasma resid") + 
  labs(subtitle =  "Cd138+ Igha+ Blimp-1+ BCMA+ Igj+",caption = sprintf("Ncells: %s",sum(data.im$plasm.resid.4=="Pure")))

wrap_plots(list(p4.wt,p4.im))
```


