---
title: "Exploring Bcell markers in Amezcua 2019 dataset"
author: "A. Berenstein"
date: "22/04/2022"
output:
  rmdformats::readthedown:
    self-contained: true
    highlight: haddock
    thumbnails: false
    css: styles.css
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file, encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'explore_cluster_signatures_v05172022.html'))})
---


```{r, message=F, warning=F,results=F}
renv::activate()
renv::restore()

library(Seurat)
library(dplyr)
library(UCell)
library(scGate)
library(ggplot2)
library(patchwork)
library(SeuratDisk)
library(ProjecTILs)
library(glmGamPoi)
source("./utils.R")
```



# Load Bcell-filtered dataset
```{r,warning=F, collapse =T,message =F}
ndim = 30; seed = 31416
panBellFile <-  "~/Dropbox/Amezcua/processed_data/PanBcellSubset.rds"  # Subseted Pan Bcell FIle with recomputed UMAP
data <- readRDS(panBellFile)
data%>%dim
data <- data%>%subset(panBcell2HR == "Pure")
data%>%dim
data$condition <- ifelse(data$orig.ident == "DS-017-WT","wt","im") 
data.bkp <- data
data.wt <- data%>%subset(orig.ident == "DS-017-WT")
data.im <- data%>%subset(orig.ident != "DS-017-WT")
```

Visualize datasets by condition
```{r,fig.width=6,fig.height=3,warning=F, collapse =T,message =F}
fig2 <- DimPlot(data,group.by = "orig.ident",split.by = "orig.ident")
ggsave("plots/fig2_panBcell.pdf",plot = fig2,width = 12, height = 4)
ggsave("plots/fig2_panBcell.png",plot = fig2,width = 12, height = 4)
ggsave("plots/fig2_panBcell.tiff",plot = fig2,width = 12, height = 4)
fig2

```

# Preprocessing

**We will preprocess separately WT and IM subsets**

## Cell cycle scoring and regression 

The idea here is to mitigate the effects of cell cycle heterogeneity by calculating cell cycle phase scores based on canonical markers, and regressing these out of the data during pre-processing (see [SeuratVignette cell cycle](https://satijalab.org/seurat/articles/cell_cycle_vignette.html) for details)

```{r,warning=F, collapse =T,message =F, results='hide'}
# extract cell cycle genes from Seurat 
s.genes <- cc.genes$s.genes%>%hs.2.mm(.)
g2m.genes <- cc.genes$g2m.genes%>%hs.2.mm(.)
```


```{r,warning=F, collapse =T,message =F, results='hide'}
# WT
set.seed(seed)
data.wt <- CellCycleScoring(data.wt, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.wt$CC.Difference <- data.wt$S.Score - data.wt$G2M.Score
data.wt <- SCTransform(data.wt, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE)
data.wt <- RunPCA(data.wt, features = data.wt@assays$RNA@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.wt <- RunUMAP(data.wt, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)


```



```{r,warning=F, collapse =T,message =F,results='hide'}
# IM
set.seed(seed)

data.im <- CellCycleScoring(data.im, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
data.im$CC.Difference <- data.im$S.Score - data.im$G2M.Score

# work as usual (only variable genes)
data.im <- SCTransform(data.im, method = "glmGamPoi", vars.to.regress = c("percent.mito","CC.Difference"), verbose = FALSE)
data.im <- RunPCA(data.im, features = data.im@assays$RNA@var.features,
                ndims.print = 1:5, nfeatures.print = 5)
data.im <- RunUMAP(data.im, reduction = "pca", dims = 1:ndim, seed.use=seed, n.neighbors = 20)


```





# Unsupervised analysis
```{r,fig.height=4, fig.width= 13,warning=F, collapse =T,message =F, results='hide'}
set.seed(seed)
data.im <- FindNeighbors(data.im,assay  = "SCT", dims = 1:20)
data.im <- FindClusters(data.im, resolution = 0.7)
data.im <- FindClusters(data.im, resolution = 0.5)
data.im <- FindClusters(data.im, resolution = 0.4)
data.im <- FindClusters(data.im, resolution = 0.3)


set.seed(seed)
data.wt <- FindNeighbors(data.wt,assay  = "SCT", dims = 1:20)
#data.wt <- FindClusters(data.wt, resolution = 2)
data.wt <- FindClusters(data.wt, resolution = 1.5)
data.wt <- FindClusters(data.wt, resolution = 1)
data.wt <- FindClusters(data.wt, resolution = 0.9)
data.wt <- FindClusters(data.wt, resolution = 0.8)
data.wt <- FindClusters(data.wt, resolution = 0.75)

data.wt <- FindClusters(data.wt, resolution = 0.7)

data.wt <- FindClusters(data.wt, resolution = 0.6)
data.wt <- FindClusters(data.wt, resolution = 0.65)

data.wt <- FindClusters(data.wt, resolution = 0.5)
data.wt <- FindClusters(data.wt, resolution = 0.4)
data.wt <- FindClusters(data.wt, resolution = 0.3)

a <- DimPlot(data.wt,group.by = "SCT_snn_res.0.8",label = T)
b <- DimPlot(data.im ,group.by = "SCT_snn_res.0.4",label = T)

a + b

```

## Explore and remove outliers
```{r,eval =F, results=F,echo =F}
#(data.im%>%subset(SCT_snn_res.0.3 == 1))$nCount_RNA%>%hist(breaks = 25)  # It seems like a bimodal distribution and a good split point could be around ncount = 1200
#data.im$nCount_RNA%>%quantile(0.95)
#data.im$nCount_RNA%>%quantile(0.9)
(data.im%>%subset(SCT_snn_res.0.3 == 1))$nFeature_RNA%>%hist(breaks = 25)  # It seems like a bimodal distribution and a good split point could be around ncount = 1200

```


```{r,warning=F, collapse =T,message =F}
## Remove outliers WT
q = 0.95
thr.wt = quantile(data.wt$nCount_RNA,q)
thr.wt.feat = quantile(data.wt$nFeature_RNA,q)
outl.wt <- data.wt%>%subset(nCount_RNA>=thr.wt | nFeature_RNA>=thr.wt.feat)
clean.wt <- data.wt%>%subset(nCount_RNA<thr.wt & nFeature_RNA<thr.wt.feat)

## Remove outliers IM
q = 0.95
thr.im = quantile(data.im$nCount_RNA,q)
thr.im.feat = quantile(data.im$nFeature_RNA,q)
outl.im <- data.im%>%subset(nCount_RNA>=thr.im | nFeature_RNA>=thr.im.feat)
clean.im <- data.im%>%subset(nCount_RNA<thr.im & nFeature_RNA<thr.im.feat)

```

```{r}
clean.wt
clean.im
```


```{r,warning=F, collapse =T,message =F}
thr.im; thr.im.feat
outl.im$SCT_snn_res.0.4%>%table()
data.im$SCT_snn_res.0.4%>%table()
```


```{r,eval =F,echo =F, results=F,collapse=T}
thr.wt; thr.wt.feat
outl.wt$SCT_snn_res.0.8%>%table()
data.wt$SCT_snn_res.0.8%>%table()

thr.im; thr.im.feat
outl.im$SCT_snn_res.0.4%>%table()
data.im$SCT_snn_res.0.4%>%table()

```


```{r,eval =F,fig.width= 8,fig.height=4}
aa <- DimPlot(outl.wt ,group.by = "SCT_snn_res.0.8",label = T) + ggtitle("wt outliers q=0.95; res = 0.8")
bb <- DimPlot(outl.im ,group.by = "SCT_snn_res.0.4",label = T) + ggtitle("IM outliers q = 0.95; res = 0.4")
aa + bb
```


```{r,eval =F,fig.width= 8,fig.height=4}

aa <- DimPlot(clean.wt ,group.by = "SCT_snn_res.0.8",label = T) 
bb <- DimPlot(clean.im ,group.by = "SCT_snn_res.0.4",label = T) 
fig3a <- aa 
fig3b <- bb

ggsave("plots/fig3a_clustering_wt_im.pdf",plot = fig3a,width = 4.5, height = 4) 
ggsave("plots/fig3a_clustering_wt_im.png",plot = fig3a,width = 4.5, height = 4)
ggsave("plots/fig3a_clustering_wt_im.tiff",plot = fig3a,width = 4.5, height = 4)

ggsave("plots/fig3b_clustering_wt_im.pdf",plot = fig3b,width = 4.5, height = 4)
ggsave("plots/fig3b_clustering_wt_im.png",plot = fig3b,width = 4.5, height = 4)
ggsave("plots/fig3b_clustering_wt_im.tiff",plot = fig3b,width = 4.5, height = 4)

aa + bb

```

```{r}
write.table(clean.im$SCT_snn_res.0.4,"~/Dropbox/Amezcua/tissue_resident_cells/reports/old_im_clusters_0_to_7.tsv",sep = "\t",quote = F,col.names = F)
write.table(clean.wt$SCT_snn_res.0.8,"~/Dropbox/Amezcua/tissue_resident_cells/reports/old_wt_clusters_0_to_3.tsv",sep = "\t",quote = F,col.names = F)

```


```{r}
original_clustes = read.table("~/Dropbox/Amezcua/tissue_resident_cells/reports/original_seurat_cluster.tsv")
original_clustes$id <- original_clustes%>%rownames()
im_clusters <- clean.im@meta.data%>%select("SCT_snn_res.0.4")
im_clusters$id <- im_clusters%>%rownames()

im_clusters = merge(im_clusters,original_clustes,by.x = "id",by.y = "id",all.x = T)
relation_ids_im = table(im_clusters$SCT_snn_res.0.4,im_clusters$seurat_clusters)
write.table(relation_ids_im,file = "~/Dropbox/Amezcua/tissue_resident_cells/reports/relation_cluster_ids_ims.tsv",sep = "\t",col.names = T,row.names =T,quote =F)


wt_clusters <- clean.wt@meta.data%>%select("SCT_snn_res.0.8")
wt_clusters$id <- wt_clusters%>%rownames()

wt_clusters = merge(wt_clusters,original_clustes,by.x = "id",by.y = "id",all.x = T)

relation_ids_wt <- table(wt_clusters$SCT_snn_res.0.8,wt_clusters$seurat_clusters)
write.table(relation_ids_wt,file = "~/Dropbox/Amezcua/tissue_resident_cells/reports/relation_cluster_ids_wt.tsv",sep = "\t",col.names = T,row.names =T,quote =F)


```


```{r}
data.wt <- clean.wt
data.im <- clean.im
```

# Marker analysis

```{r,fig.height= 20,fig.width=10,warning=F, collapse =T,message =F}
Idents(data.wt) <- "SCT_snn_res.0.8"
discrim.wt <- FindAllMarkers(data.wt, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.wt <- discrim.wt%>%subset(p_val_adj < 0.05)
gene.wt <- discrim.wt$gene
d <- DoHeatmap(data.wt,features = gene.wt)

Idents(data.im) <- "SCT_snn_res.0.4"
discrim.im <- FindAllMarkers(data.im,assay = "SCT", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.im.uncorrected <- discrim.im
discrim.im <- discrim.im%>%subset(p_val_adj < 0.05)

gene.im <- discrim.im$gene

e <- DoHeatmap(data.im,features = gene.im,size = 3,lines.width = 35)

# show WT heatmap
d 

# show IM heatmap
e 
```


```{r,eval = F}
ggsave(d,file = "~/repos/ajberenstein/tissueResidentCells/docs/heatmap_fdr005.wt.png",height = 20,width = 10, dpi = 300)
ggsave(d,file = "~/repos/ajberenstein/tissueResidentCells/docs/heatmap_fdr005.wt.pdf",height = 20,width = 10,dpi = 300)
write.table(discrim.wt,file = "~/repos/ajberenstein/tissueResidentCells/docs/makers_fdr005.wt.tsv",sep = "\t",quote =F,row.names =F)

ggsave(e,file = "~/repos/ajberenstein/tissueResidentCells/docs/heatmap_fdr005.im.png",height = 20,width = 10,dpi = 300)
ggsave(e,file = "~/repos/ajberenstein/tissueResidentCells/docs/heatmap_fdr005.im.pdf",height = 20,width = 10,dpi = 300)
write.table(discrim.im,file = "~/repos/ajberenstein/tissueResidentCells/docs/makers_fdr005.im.tsv",sep = "\t",quote =F,row.names =F)
```


```{r,eval = F}
# 
discrim.wt.down <- FindAllMarkers(data.wt, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.wt.down <- discrim.wt.down%>%subset(p_val_adj < 0.05)
write.table(discrim.wt.down,file = "~/repos/ajberenstein/tissueResidentCells/docs/makers_fdr005.wt_include_downReg.tsv",sep = "\t",quote =F,row.names =F)


discrim.im.down <- FindAllMarkers(data.im, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.15)
discrim.im.down <- discrim.im.down%>%subset(p_val_adj < 0.05)
write.table(discrim.im.down,file = "~/repos/ajberenstein/tissueResidentCells/docs/makers_fdr005.im_include_downReg.tsv",sep = "\t",quote =F,row.names =F)


```



```{r,eval=F,echo=F,results=F}
discrim.im$cluster%>%table()
discrim.im%>%head((discrim.im$cluster%>%table())["0"])
discrim.im%>%subset(cluster==5)
```


# DimPlot using Heatmap genes

Plot again clusters in WT and IM datasets


```{r,fig.height=4, fig.width= 10,warning=F, collapse =T,message =F, results='hide'}
a + b
```


WT dataset
```{r,fig.width= 15,fig.height= 50,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = discrim.wt$gene,ncol = 5)
```

IM dataset
```{r,fig.width= 20,fig.height=70,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = intersect(discrim.im$gene,data.im%>%rownames()),ncol = 6)
```


# Explore predefined markers 

We will use here those genes previously defined in our molecular signature spreadsheet file

```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
features1 = c("Ighm","Ighd1-1","Igha","Ighg1","Cd19","Ms4a1","Fcer2a","Cd24a","Cd38","Pdcd1lg2","Nt5e","Entpd1","Cd80","Cd86","Cd93","Cd69","Itgal","Cd44","Cd28","Sdc1",	"Prdm1","Tnfrsf17","Sell","Igj")
features2 <- c("Fas","Bcl6","Cd40","Cxcr5","Cxcr4")
```

```{r,fig.height= 8,fig.width=8,warning=F, collapse =T,message =F}
h1 <- DoHeatmap(data.wt,features = features1)
h2 <- DoHeatmap(data.im,features = features1,size = 3,lines.width = 35)

h1
h2
```


WT dataset


```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = features1, ncol = 5)
```

IM dataset
```{r,fig.height=15,fig.width=15,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = features1, ncol = 5)
```


WT dataset (new genes)
```{r,fig.height=7,fig.width=12,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = features2, ncol = 3)
```

IM dataset (new genes)
```{r,fig.height=7,fig.width=12,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = features2, ncol = 3)
```


# Explore markers of resident plasmatic cells (IgA+)
WT dataset
```{r,fig.width= 12,fig.height=8,warning=F, collapse =T,message =F}
FeaturePlot(data.wt,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```

IM dataset
```{r,fig.width= 13,fig.height=7,warning=F, collapse =T,message =F}
FeaturePlot(data.im,features = c("Sdc1","Igha","Prdm1","Tnfrsf17","Cd28","Igj","Cd93"),ncol = 4)
```

Include scGate model
```{r,fig.height=4,fig.width=8,warning=F, collapse =T,message =F}
plasm.resid.4 <- scGate::gating_model(name = "plasmaticas_residentes", signature = "Sdc1;Igha;Prdm1;Tnfrsf17;Igj")
data.im <- scGate(data = data.im, model = plasm.resid.4, output.col.name = "plasm.resid.4")
p4.im <- DimPlot(data.im, cols = c(list(Impure = "gray", Pure = "green")), group.by = "plasm.resid.4") + theme(aspect.ratio = 1) + ggtitle("plasma resid") + 
  labs(subtitle =  "Cd138+ Igha+ Blimp-1+ BCMA+ Igj+",caption = sprintf("Ncells: %s",sum(data.im$plasm.resid.4=="Pure")))

data.wt <- scGate(data = data.wt, model = plasm.resid.4, output.col.name = "plasm.resid.4")
p4.wt <- DimPlot(data.wt, cols = c(list(Impure = "gray", Pure = "green")), group.by = "plasm.resid.4") + theme(aspect.ratio = 1) + ggtitle("plasma resid") + 
  labs(subtitle =  "Cd138+ Igha+ Blimp-1+ BCMA+ Igj+",caption = sprintf("Ncells: %s",sum(data.im$plasm.resid.4=="Pure")))

wrap_plots(list(p4.wt,p4.im))
```


